---
title: "R Notebook"
output: html_notebook
---


```{r}
# required libraries

library("RStoolbox")
library("raster")
library("rgdal")
library("ggplot2")
library("SDMTools")

```


```{r}
# load tif using stack function instead raster

TTC08335_modified.stack <- stack("../6_qgis/input/TTC08335_modified.tif")
class(TTC08335_modified.stack)

```


```{r}
# load parcela 1
library(rgeos)

parcelas <- readOGR("../6_qgis/input/alamala.kml", "alamala")[0]
nrowdim <- dim(parcelas@data)
parcelas@data$id <- c(rep(1:nrowdim))
parcela2 <- subset(parcelas, parcelas@data$id %in% c(3,4))
p2.centroid.df <- as.data.frame(gCentroid(parcela2, byid = TRUE)) # extract to label parcels
# p2.21  <- gCentroid(parcela2, byid = TRUE)[1,]
# p2.22  <- gCentroid(parcela2, byid = TRUE)[2,]

class(parcela2)
plot(parcela2)
```


```{r}
parcela2.df <- fortify(parcela2) # to plot with ggplot
head(parcela2.df)
```


```{r}


# Set all pixels to NA, where bands are 0 (remove black background)
# Check if results are affected
# instead use crop and mask together (ver más adelante)

TTC08335_modified.stack[TTC08335_modified.stack[,] == 0] <- NA

parcela2.df

# plot scene using ggRGB (from ggplot and RStoolbox)
ggRGB(TTC08335_modified.stack, r = 1, g = 2, b = 3, maxpixels = 2e+05, stretch="none", geom_raster = TRUE) + 
    geom_path(data = parcela2.df, aes(x = long, y = lat, group = group), size = 1, col="#fbae3b") +
    geom_text(data = as.data.frame(p2.centroid.df), aes(label = c("2.1", "2.2") ,y = y, x = x), colour = "white")+ 
    coord_equal() +
    theme_bw()
```



```{r}
ggsave("figures/parcela2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```



Extraemos las parcelas

```{r}
# crop and mask whole area
 TTC08335_modified.stack_p2 <- crop(mask(TTC08335_modified.stack, parcela2),parcela2)

```



```{r}

# crop parcela 2.1

parcela21 <- subset(parcelas, parcelas@data$id %in% c(3))
TTC08335_modified.stack_p21 <- crop(mask(TTC08335_modified.stack, parcela21), parcela21)
 
# crop parcela 2.2
parcela22 <- subset(parcelas, parcelas@data$id %in% c(4))

TTC08335_modified.stack_p22 <- crop(mask(TTC08335_modified.stack, parcela22), parcela22)

```


Plot parcels

```{r}

ggRGB(TTC08335_modified.stack_p2, r = 1, g = 2, b = 3) + 
    geom_path(data = parcela2.df, aes(x = long, y = lat, group = group), size = 1, col="#fbae3b") +
    theme_bw()
```


Plot parcel 2.1

```{r}
p21 <- subset(parcela2.df, id == 2)

ggRGB(TTC08335_modified.stack_p21, r = 1, g = 2, b = 3) + 
    geom_path(data = p21, aes(x = long, y = lat, group = group), size = 1, col="#fbae3b") +
    labs(x="", y="", title="Parcela 2.1") +
   # coord_equal(ylim = c(min(p11$lat), max(p11$lat)), xlim= c(min(p11$long), max(p11$long))) +
    theme_bw() + 
    theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))  # make title bold and add spac

```
```{r}
ggsave("figures/parcela2_1.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


Plot parcel 2.2

```{r}

p22 <- subset(parcela2.df, id == 4)


ggRGB(TTC08335_modified.stack_p22, r = 1, g = 2, b = 3) + 
    geom_path(data = p22, aes(x = long, y = lat, group = group), size = 1, col="#fbae3b") +
    labs(x="", y="", title="Parcela 2.2")+
    #coord_equal(ylim = c(min(p12$lat), max(p12$lat)), xlim= c(min(p22$long), max(p22$long))) +
    theme_bw() + 
    theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))  # make title bold and add spac

```
```{r}
ggsave("figures/parcela2_2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```



Calculate VIs

```{r}

p21.VIs <- spectralIndices(TTC08335_modified.stack_p21, green=3, red=2,  nir =1, indices=c("NDVI", "MSAVI2", "GNDVI"))
breaks <- seq(0, 1, by=0.01)

plot(p21.VIs)
```


Plot VI one by one


NDVI Parcela 1.1

```{r}
cols <- colorRampPalette(c("red", "yellow", "lightgreen"))(length(breaks)-1)
ggR(p21.VIs$NDVI, geom_raster = TRUE) +
  labs(x="", y="", title= "Parcela 2.1") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))

```


And save

```{r}
ggsave("figures/parcela2_1_NDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


MSAVI2 Parcela 2.1

```{r}
cols <- colorRampPalette(c("red", "yellow", "lightgreen"))(length(breaks)-1)
ggR(p21.VIs$MSAVI2, geom_raster = TRUE) +
  labs(x="", y="", title= "Parcela 2.1") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))

```
And save

```{r}
ggsave("figures/parcela2_1_MSAVI2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


GNDVI Parcela 1.1

```{r}
cols <- colorRampPalette(c("red", "yellow", "lightgreen"))(length(breaks)-1)
ggR(p21.VIs$GNDVI, geom_raster = TRUE) +
  labs(x="", y="", title= "Parcela 2.1") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))

```
```{r}
ggsave("figures/parcela2_1_GNDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


Again with p1.2

```{r}
p22.VIs <- spectralIndices(TTC08335_modified.stack_p22, green=3, red=2,  nir =1, indices=c("NDVI", "MSAVI2", "GNDVI"))
breaks <- seq(0, 1, by=0.01)

plot(p22.VIs)
```

```{r}
ggR(p22.VIs$NDVI, geom_raster = TRUE) +
  labs(x="", y="", title="Parcela 2.2") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))  # make title bold and add spac

```


```{r}
ggsave("figures/parcela2_2_NDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

MSAVI2 Parcela 2.2

```{r}
ggR(p22.VIs$MSAVI2, geom_raster = TRUE) +
  labs(x="", y="", title="Parcela 2.2") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))  # make title bold and add spac

```

Save

```{r}
ggsave("figures/parcela2_2_MSAVI2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

x`
GNDVI Parcela 2.2

```{r}
ggR(p22.VIs$GNDVI, geom_raster = TRUE) +
  labs(x="", y="", title="Parcela 2.2") +
  scale_fill_gradientn(colours=cols,  na.value=NA) + 
  theme_bw() +
  theme(plot.title = element_text(lineheight=.8, face="bold", vjust=1, hjust = 0.5))  # make title bold and add spac

```


Save

```{r}
ggsave("figures/parcela2_2_GNDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


### Exploratory data analysis of Vegetation Indices


Bind indices in a whole dataframe

```{r}
library(dplyr)
library(plyr) # Tools for Splitting, Applying and Combining Data

raster_to_df <- function(x) {
  stack(as.data.frame(x))
} # convert raster to dataframe


l<- list(p21 = p21.VIs, p22 = p22.VIs)

l.df <- lapply(X = l, FUN = raster_to_df) # list of data frames


l.df.VIs <- ldply(l.df ,rbind) # Split list, apply function, and return results in a data frame.


```



Plot NDVI box-plot

```{r}
l.df.VIs.NDVI <-  subset(l.df.VIs, ind == "NDVI" )
ggplot(l.df.VIs.NDVI) + 
  geom_boxplot(aes(x = .id, y = values, colour=.id)) +
  facet_grid(. ~ ind) +
  theme_bw()
```



```{r}
ggsave("figures/boxplot_p21_p22_NDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```



Plot MSAVI2 box-plot


```{r}
l.df.VIs.MSAVI2 <-  subset(l.df.VIs, ind == "MSAVI2" )
ggplot(l.df.VIs.MSAVI2) + 
  geom_boxplot(aes(x = .id, y = values, colour=.id)) +
  facet_grid(. ~ ind) +
  theme_bw()
```

```{r}
ggsave("figures/boxplot_p21_p22_MSAVI2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```


Plot GNDVI box-plot


```{r}
l.df.VIs.GNDVI <-  subset(l.df.VIs, ind == "GNDVI" )
ggplot(l.df.VIs.GNDVI) + 
  geom_boxplot(aes(x = .id, y = values, colour=.id)) +
  facet_grid(. ~ ind) +
  theme_bw()
```
```{r}
ggsave("figures/boxplot_p21_p22_GNDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

Plot NDVI, MSAVI2, GNDVI histogram


```{r}
l.df.VIs.NDVI$title <- "NDVI" # fake


ggplot(l.df.VIs.NDVI, aes(x = values, colour=.id)) + 
  geom_freqpoly(aes( y=(..count..)/sum(..count..)), binwidth = 0.005) +
  facet_wrap(~title) +
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + 
  theme_bw()
```

```{r}
ggsave("figures/histo_p21_p22_NDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

```{r}
l.df.VIs.GNDVI$title <- "MSAVI2" # fake


ggplot(l.df.VIs.GNDVI, aes(x = values, colour=.id)) + 
  geom_freqpoly(aes( y=(..count..)/sum(..count..)), binwidth = 0.005) +
  facet_wrap(~title) +
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + 
  theme_bw()
```

```{r}
ggsave("figures/histo_p21_p22_MSAVI2.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

```{r}
l.df.VIs.GNDVI$title <- "GNDVI" # fake


ggplot(l.df.VIs.GNDVI, aes(x = values, colour=.id)) + 
  geom_freqpoly(aes( y=(..count..)/sum(..count..)), binwidth = 0.005) +
  facet_wrap(~title) +
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + 
  theme_bw()
```

```{r}
ggsave("figures/histo_p21_p22_GNDVI.png", 
 plot = last_plot(), # or give ggplot object name as in myPlot,
 width = 5, height = 5, 
 units = "in", # other options c("in", "cm", "mm"), 
 dpi = 300)
```

# Analysis


```{r}
library('dplyr')
l.df.VIs %>% group_by(ind) %>% summarise_at("values", funs(mean, max, sd), na.rm = TRUE)

```

```{r}
library('dplyr')
l.df.VIs %>% group_by(ind, .id) %>% summarise_at("values", funs(mean, max, sd), na.rm = TRUE)
```


```{r}
saveRDS(l.df.VIs, file = "VIP2.rds")
```


```{r}
# save p2 layers
writeRaster(stack(p21.VIs), paste("p21_", names(p21.VIs), sep = ''), bylayer=TRUE, format='GTiff')
writeRaster(stack(p22.VIs), paste("p22_", names(p22.VIs), sep = ''), bylayer=TRUE, format='GTiff')

```